Understanding GDP Relations to Crime in US Cities

### 2017-03-08 resource/index.Rmd
more work on resource

### What Are The Datasets?
()
The two datasets that we used in this anaylsis are one for crime based data on U.S. cities over a 5 year time period (2010-2015) and one for the GDP for metropolitan areas in the U.S. over a 5 year time period (2010-2015).

@@

While it makes sense that as Populuation increases the number of crimes increase, but we believed that just that information is not what should be the focus when looking at trends in crime.  Instead we created a value that we named *Crime Rate* that is the total crimes committed in a city divided by the population.  This new value is the total number of crimes per person within a city, or what we refer to as the crime rate.  _Note:_ This is not to be confused with the more common term "crime rate" that refers to the number of crimes committed per 100,000 people. 
<br> 
<br> 
So what we did is we took the newly made variable *Crime Rate* and compared it with the population of the different cities and their Gross Domestic Product per Person. If we break all that down here are how the different data points relate to each other
``` {r echo=FALSE, message=FALSE, warning=FALSE}
colnames(time.poisson.data)[4] <- "Year"

@@

What this breaks down to is that for every one dollar that GDP per person increase the Crime rate increase by 0.0000004938 and for every person added to the population crme rate decreases by 0.0000000527.  Theses numbers seem incredibly small but when you are talking about millions of dollars in GDP and thousands of people it becomes increasingly important.

###

>>>>

### What Are The Datasets?

The two datasets that we used in this anaylsis are one for crime based data on U.S. cities over a 5 year time period (2010-2015) and one for the GDP for metropolitan areas in the U.S. over a 5 year time period (2010-2015).

@@

While it makes sense that as Populuation increases the number of crimes increase, but we believed that just that information is not what should be the focus when looking at trends in crime.  Instead we created a value that we named *Crime Rate* that is the total crimes committed in a city divided by the population.  This new value is the total number of crimes per person within a city, or what we refer to as the crime rate.  _Note:_ This is not to be confused with the more common term "crime rate" that refers to the number of crimes committed per 100,000 people. 
<br> 
<br>

###Crime Rate with GDP and Population

So what we did is we took the newly made variable *Crime Rate* and compared it with the population of the different cities and their Gross Domestic Product per Person. If we break all that down here are how the different data points relate to each other
``` {r echo=FALSE, message=FALSE, warning=FALSE}
colnames(time.poisson.data)[4] <- "Year"

@@

What this breaks down to is that for every one dollar that GDP per person increase the Crime rate increase by 0.0000004938 and for every person added to the population crme rate decreases by 0.0000000527.  Theses numbers seem incredibly small but when you are talking about millions of dollars in GDP and thousands of people it becomes increasingly important.  After computing this equation it was still found to not be statistically signifigant, meaning that the GDP per Person and the Population had no real impact on the Crime Rate.

### Crime Rate Change

After looking at just the crime rate over the 5 year time period we created a new variable which was the difference in crime rate, GDP per person, and Population change between the years.  At this point we had started to work more with countinous data than count data, so we decided to use a standard linear regression to find the equation and signifigance of the data.

``` {r echo=FALSE, message=FALSE, warnings=FALSE}
rate.crime.vs.pop <- ggplot(tmp, aes(x=population_change_value, y=total_crime_rate_change_value)) +
                  geom_point(shape=1) +
                  stat_smooth(method = "lm", col = "red") +
                  ggtitle("Population and Total Crime Rate Change") +
                  labs(x ="Population", y="Total Crime Rate Change")
rate.crime.vs.pop
```

``` {r echo=FALSE, message=FALSE, warnings=FALSE}
rate.crime.vs.pop <- ggplot(tmp, aes(x=gdp_rate_change_value, y=total_crime_rate_change_value)) +
                  geom_point(shape=1) +
                  stat_smooth(method = "lm", col = "red") +
                  ggtitle("GDP Rate Change and Total Crime Rate Change") +
                  labs(x ="GDP Rate Change", y="Total Crime Rate Change")
rate.crime.vs.pop
```

This is where the data got very interesting.  As we can see there is a very signifigant change on the Difference in the rate of crime based on the population in a negative fashion.  This means that every time the population difference increase the rate of crime decreases.  But on the other side of thins the GDP change actually increases the Crime change.  Both of these factors were statistically signifigant, meaning the both affect the difference in the crime rate.

### What this all means.
